---
title: "Insurance"
output:
  word_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown

The read.csv is used to import the data, and file.choose helps one bypass the troubles of setting working directory.

Let's dig into the insurance data set.

```{r}
insurance<-read.csv(file.choose(),header=T)
```



```{r}
library(tidyverse)
library(ggplot2)
library(scales)
library(extrafont)
```

Some times I pick and choose between head and View. But I use them to have a quick preview about the data.

```{r}
View(insurance)

head(insurance, 10)
```


Let's see how R imported the data. A close look at data types

```{r}
str(insurance)
```
Now some data Viz to explore what our data looks like.

First I'll like to see the spread age data by sex. Do we have older males or females.  
```{r}
ggplot(insurance)+
  geom_boxplot(aes(sex,  age))+theme_classic()


```

From the box plot the data the age data looks spread evenly between male and female, I'll take a closer look using the dotplot geom

```{r}
ggplot(insurance)+
  geom_dotplot(aes(sex,  age), binaxis="y", dotsize=1, binwidth=0.5, stackdir="center")+theme_classic()
```
The dotplot shoes us something interesting in the ~20 age group. Younger study population?

Time to visualize the age distribution by region and smoking

```{r}
ggplot(insurance)+
  geom_boxplot(aes(sex,  age, color=sex))+theme_classic() +facet_grid(~region)


```

```{r}
ggplot(insurance)+
  geom_boxplot(aes(sex,  age, color=smoker))+theme_classic() +facet_grid(~smoker)
```

Some summary analysis to view the charges by region grouping the data by sex.

First I'll create a summary data to view the average charges by region

```{r}
charges_by_region <- insurance %>% 
  group_by(region) %>%
  summarise(Avg_region_charges=mean(charges)) %>% ungroup

charges_by_region
```

Time to visualize 

```{r}
charges_by_region %>% 
 ggplot(aes(reorder(region,Avg_region_charges),Avg_region_charges,fill=region))+
  geom_col(show.legend = FALSE,color="black")+
geom_text(aes(label=comma(Avg_region_charges)),size=3,hjust=1,color="black")+
  scale_fill_brewer(palette = "Paired")+
  coord_flip()+
  theme_classic()+
  labs(title = "Average Insurance Charges by Region",x="region",y= "Charges") 
```



Next We visualize Charges by sex

```{r}
charges_by_sex <- insurance %>% 
  group_by(sex) %>%
  summarise(Avg_region_charges_sex=mean(charges)) %>% ungroup

charges_by_sex
```

Males are charges more? Does our data have any answer to this? Do we have more males in the southeast? Does smoking status or BMI affect charges? 

Lets get on some visualizations to answer these questions.

```{r}
ggplot(insurance)+
  geom_point(aes(bmi,charges, color=charges, size=charges))+
  scale_color_gradient(low="orange", high="purple")+
  theme_classic()+
  labs(title = "Charges by BMI",x="BMI",y= "Charges") 
 
```
Lets View this data by Region and Sex

```{r}
ggplot(insurance)+
  geom_point(aes(bmi,charges, color=charges))+
  scale_color_gradient(low="orange", high="purple")+
  theme_classic()+
  labs(title = "Charges by BMI",x="BMI",y= "Charges") + facet_grid(.~region+sex)

```

FIrst we see that most of the charges are below 20000, also there is no defined relationship between charges and bmi. 
Can age be the answer?

```{r}
ggplot(insurance)+
  geom_point(aes(age,charges, color=charges))+
  scale_color_gradient(low="orange", high="purple")+
  theme_classic(base_size=15)+
  labs(title = "Charges by AGE",x="AGE",y= "Charges") 
```




```{r}
lmBMI = lm(charges~bmi, data = insurance)
summary(lmBMI)
```
The BMi is a good addition to the model P = 2.46e-13 ***, However the R square shows that this model is a poor fit for the data. 

Now lets try charges by age


```{r}
lmage = lm(charges~age, data = insurance)
summary(lmage)
```


Similarly the age is a good predictor for the model as the P value is < 2e-16. However this model explains only 9 percent of the variation in the model. 

Human studies typically have low R2 because of unpredictability, but will 4 percent or 9 percent explained variation pass ? I don't don't so.

I'll try a multiple regression before I come to initial conclusions. 

```{r}
lmagebmi = lm(charges~age + bmi, data = insurance)
summary(lmagebmi)
```

VIola! both age and bmi are good predictors for charges, however their R2 is only 12 percent. Is this model useable? Yes!First this data set might be a subset of a large data set and the sample size has been reduced, also, this might be a streamlined data set, where other variables that will bolster the model have been removed.

I added children to the regression and the R2 increased to 13 percent.

What is your take on this interpretation. 



Final phase of working with this data, I want to answer the question " Does region, smoker or sex" have any effect on charges


The first thing I do is ensure I have my factors in place and check the data types after chnaging them.

```{r}
insurance$sex=as.factor(insurance$sex)

insurance$smoker=as.factor(insurance$smoker)

insurance$region=as.factor(insurance$region)

str(insurance)
```
```{r}
insaov=aov(charges~(smoker*sex)+region,data=insurance)
anova(insaov)
```
Now it all makes sense!!!!!!! Charges is statistically similar for sex and region!!! Though I blocked by region, which is a random factor. I didn't expect a difference in charges by region.

Now lets do the final visualizations to see what the data looks like by smoking....

```{r}
charges_by_smoke <- insurance %>% 
  group_by(smoker) %>%
  summarise(Avg_smoke_charges=mean(charges)) %>% ungroup

charges_by_smoke
```

```{r}
charges_by_smoke %>% 
 ggplot(aes(reorder(smoker,Avg_smoke_charges),Avg_smoke_charges,fill=smoker))+
  geom_col(show.legend = FALSE,color="black")+
geom_text(aes(label=comma(Avg_smoke_charges)),size=5,hjust=1,color="black")+
  scale_fill_brewer(palette = "Paired")+
  coord_flip()+
  theme_classic()+
  labs(title = "Average Insurance Charges by Smoking",x="Smoker",y= "Charges") 
```
The End on today's episode of telling stories with datasets. 
